Spreading of a Facebook Application

### 2011-04-15 paper.tex
changes up to 4.2

Results show that participants who connect the network of existing
users to a new one often achieved a relatively high score in the quiz. New
participants often had strong connections to the network of people who
already participated. Indeed, the number of friends did not seem to make
users more or less likely to gather new participants for the quiz.
\end{abstract}

@@

\section{Introduction}
\label{sec:introduction}
Founded in 2004, Facebook\footnote{www.facebook.com} has
become a social network currently used by more than 600 million people
all over the world\cite{facebook500}. Since 2007, Facebook provides an API permitting
third party developers to embed their own applications in
Facebook\footnote{http://developers.facebook.com/blog/archive}.
Subsequent to the F8 conference in April, 2010, Facebook simplified
access to its social graph by allowing
external websites for integration in their social graph through Open
Graph Protocol\cite{facebookStats} .

Due to integration in Facebook's social graph, third party
applications not only spare user authentication procedures, but can also access
extracts of Facebook's social graph, too. These information provided
by Facebook detail a user's friend list, basic information, and
further data to which access has to be granted by access tokens
\footnote{http://developers.facebook.com/docs/authentication/}. Thus,
applications let Facebook's social graph enlarge and make people
spend more time of Facebook. In addition, application developers
can spread their application virally through appearance in news
streams, and collect useful information about the social network of
their users, too.

@@

In the first section, basic information about Facebook's social graph,
and integration of applications is
given. In the next section, the methodology is described by detailing
the quiz and its building, as well as the tool for data evaluation. Section
\ref{sec:results} presents the results, and in section
\ref{sec:discussion} and \ref{sec:outlook}, a discussed follows. Section
\ref{sec:conclusion} concludes the paper by summing up the results,
and giving an outlook for further research, too.

\section{Foundations}
\label{sec:background}
This section details concepts from Facebook, basics about the
\ac{DOT} language, and at least, it defines the metrics the tool
explained in section \ref{sec:methodology} can output. Facebook's
social graph concept is explained in 
section \ref{sec:social-graph}, section \ref{sec:integr-appl-with}
describes how applications can be integrated with Facebook, and
section \ref{sec:auth-auth} makes clear how applications can be
integrated with Facebook.

@@

Facebook provides several features to support the development
process. \ac{sdks} exist for popular programming languages and provide
sets of pre-defined functions for communication with Facebook. For
instance, they include the O-Auth 2.0 dialog which is described in
section \ref{sec:auth-auth}. In
addition, Facebook grants observation of its own development process by
making its raodmap public, and by showcasting its buglist. In addition, a
testserver exists to help developers to analyze if their application
is going to run with planned features from Facebook. 
Facebook provides an \textit{insights}-feature which makes
statistics about the applications available. \cite{oreilly}.

\subsection{Authentication and Authorization}
\label{sec:auth-auth}
In order to authorize and authenticate, Facebook recommends usage of
O-Auth 2.0 protocol by using the OAuth Dialog included in its
\ac{sdks}. As the application analyzed in this paper and described in section 
\ref{sec:method} implements the O-Auth Dialog from the PHP-\ac{sdk},
this variant will be described in detail. It shall just be mentioned,
that Facebook also allows for manual execution of the steps to receive
access tokens.\cite{facebookDevAuth}

@@

When clients access the canvas page of the application, they are
redirected to the OAuth Dialog (running on a Facebook server) and
provided with the application's ID. Then, user authentication and
app authorization is performed, by involving the client and Facebook
only. User authentication is done by logging in on Facebook (if necessary) and
authorization of the application is done by asking the user if he/she
wants to grant the application access to some data specified by the
application with the scope parameter. One should note, that no
login data are sent to the application, thus, this is how Facebook
establishes a trust relation. 
If the user authorizes the application, an \textit{authorization code} is returned and
the user is redirected to the application, to which the
\textit{authorization code} is forwarded. During registration of the
application, an \textit{application secret} code was
created. In order to authenticate the application, the
\textit{authorization code} and the \textit{application secret} have
to be sent to Facebook's \textit{Graph API Token Endpoint}. Facebook
then returns an \textit{access token}, which can be included in the
URL by which the Graph API can be accessed. However, Facebook's
\ac{sdks} also provide functions to perform calls to the Graph API,
and in this case, the access token is included automatically.

\subsection{Graphs with DOT language}
\label{sec:graph-repr-with}
The \ac{dot} language allows for plain text description of
graphs. There are few programs that can process \ac{dot} graphs, most
of them are part of the
\textit{Graphviz}\footnote{http://www.graphviz.org}-package. 
The language allows for directed and undirected graphs, attributes can
be set and colors as well as different shapes for nodes and edges can
be used. The basic syntax is very easy, the graph shown in figure
\ref{fig:dot} has the \ac{dot}-code given in listing \ref{list:dot}.

@@

\subsection{Software Metrics}
\label{sec:metrics}
Metrics define a mapping of a `particular measurable entity to a
numerical value' \cite{Lanz06a}. Thus, they allow for analysis and summarization
of properties which are oftentimes not measurable. In addition, they
facilitate detection of `outliers in large amounts of data'
\cite{Lanz06a}. Metrics can be plotted as diagrams or charts, or they
can be used for visualizations such as graphs connecting nodes
characterized by metrics. Since humans are trained in understanding signs and
pictures, visualizations supports understanding and identifying of
`hidden aspects'. \cite{Lanz06a} In this paper, they are
employed to characterize users of the quiz. Some metrics can be output
as \ac{pdf}, and others are used to create the graphs in \ac{dot} language.

@@

\subsubsection{About The Quiz}
\label{sec:aboutQuiz}
The quiz is about dental health fact and also challenging. It can be
found at the URL \url{http://apps.facebook.com/zahniquiz}.
Each participants is asked five questions randomly selected
from a set of twenty questions. A question has four answers, of
which either one, two, three or each answer can be correct, thus
the kind of answers refers to the term \textit{multiple selection}.
A question is scored to be answered correctly, if the correct subset
was selected by the user, thus each correct and no incorrect answer
were selected. Each question answered correctly scores one, thus the
highest score is five. Figure \ref{fig:quiz} gives a screenshot of the quiz.

@@

\subsubsection{Implementation}
\label{sec:implementation}
The studied application makes use of 
Facebook's php-\ac{sdk} and JavaScript-\ac{sdk}\footnote{http://developers.facebook.com/docs/}. It requires
permissions (see \ref{sec:auth-auth}) for \textit{user$\_$birthday} and
\textit{publish$\_$stream}.
The questions and answers of the quiz are stored in a MySQL-database. PHP scripts
handle the creation of the quiz and
the Pear DB module \footnote{http://pear.php.net/} is used
for database queries.

@@

\paragraph{Application flow}
\label{sec:application-flow}
When a user accesses the canvas page of the quiz, 
the \textit{OAuth Dialog} is started automatically by the
\textit{controller}. If authorization and authentication are completed
successfully, the user's basic data as well as the birthday are stored together with
a timestamp in the \textit{User} table. Column \textit{has-used-app} is set to 1. In
addition, each user's friends list is also stored in \textup{User} and
\textit{Friends} table, and for these \textit{users}, columns
\textit{has-used-app} is set to 0.

% the quiz is created by randomly choosing
% five questions from the questions table and requiring the answers. To
% make the questions more attractive, the name of an image convenient to
% the questions topic is also stored in the questions table, and this
% image is placed on the left side of each question. Since multiple
% selection is possible, before the answers, which are also queried from
% the MySQL database, checkboxes are placed before each answer. 
Next, the quiz is created as mentioned in section
\ref{sec:implmenetation}.
After completion, its is evaluated the final score is computed. As
each answer ought to be stored and multiple selection of answers is
possible, user answers have to be stored in  a multidimensional.

@@

\subsection{Graphs detailing the spreading of the quiz}
\label{sec:distribution-quiz}
In this section, graphs illustrating how the
application spread are presented. The original user names collected
from Facebook were made anonymous
 (see section \ref{sec:data-evaluation}). Friendships are
represented by directed edges: If user A has user B in the friend list, an
arrow exists targeting user B.

>>>>

Results show that participants who connect the network of existing
users to a new one often achieved a relatively high score in the quiz. New
participants often had strong connections to the network of people who
already participated. Indeed\annSBW{passt da mal überhaupt nicht}, the number of friends did not seem to make
users more or less likely to gather new participants for the quiz.
\end{abstract}

@@

\section{Introduction}
\label{sec:introduction}
Founded in 2004, Facebook\footnote{www.facebook.com} has
become a social network currently used by more than 600 million people
all over the world\cite{facebook500}. Since 2007, Facebook provides an API permitting
third party developers to embed their own applications in
Facebook\footnote{http://developers.facebook.com/blog/archive}\annSBW{sowas kannste nicht verlinken ... wieso sind das fussnoten und keine Quellen?}.
Subsequent to the F8 conference in April, 2010, Facebook simplified
access to its social graph by allowing
external websites for integration\annSBW{falsche grammatik} in their social graph through \annSBW{the} Open
Graph Protocol\cite{facebookStats} .

Due to integration in Facebook's social graph, third party
applications not only spare user\annSBW{s from} authentication procedures, but can also access
extracts of Facebook's social graph, too. These information provided
by Facebook detail a user's friend list, basic information, and
further data to which access has to be granted by access tokens
\footnote{http://developers.facebook.com/docs/authentication/}. Thus,
applications let Facebook's social graph enlarge and make people
spend more time of Facebook. In addition, application developers
can spread their application virally through appearance in news
streams, and collect useful information about the social network of
their users, too.

@@

In the first section, basic information about Facebook's social graph,
and integration of applications is
given. In the next section, the methodology is described by detailing
the quiz and its building, as well as the tool for data evaluation. Section
\ref{sec:results} presents the results, and in section
\ref{sec:discussion} and \ref{sec:outlook}, a discussed follows \annSBW{wording}. Section
\ref{sec:conclusion} concludes the paper by summing up the results,
and giving an outlook for further research, too\annSBW{loose the too}.

\section{Foundations}
\label{sec:background}
This section details concepts from Facebook, basics about the
\ac{DOT} language, and at least, it defines the metrics the tool
explained in section \ref{sec:methodology} can output. Facebook's
social graph concept is explained in 
section \ref{sec:social-graph}, section \ref{sec:integr-appl-with}
describes how applications can be integrated with Facebook, and
section \ref{sec:auth-auth} makes clear \annSBW{no it doesn't} how applications can be
integrated with Facebook.

@@

Facebook provides several features to support the development
process. \ac{sdks} exist for popular programming languages and provide
sets of pre-defined functions for communication with Facebook. For
instance, they include the O-Auth 2.0 dialog which is described in
section \ref{sec:auth-auth}. In
addition, Facebook grants observation of its own development process by
making its raodmap \annSBW{typo} public, and by showcasting \annSBW{typo and wrong word (showcasing is positiv)} its buglist. In addition, a
testserver exists to help developers to analyze if their application
is going to run with planned features from Facebook. 
Facebook provides an \textit{insights}-feature which makes
statistics about the applications available. \cite{oreilly}.

\subsection{Authentication and Authorization}
\label{sec:auth-auth}
In order to authorize and authenticate, Facebook recommends usage of
O-Auth 2.0 protocol by using the OAuth Dialog included in its
\ac{sdks}. As the application analyzed in this paper and described in section 
\ref{sec:method} implements the O-Auth Dialog from the PHP-\ac{sdk},
this variant will be described in detail. It shall just be mentioned,\annSBW{ loose the comma}
that Facebook also allows for manual execution of the steps to receive
access tokens.\cite{facebookDevAuth}

@@

When clients access the canvas page of the application, they are
redirected to the OAuth Dialog (running on a Facebook server) and
provided with the application's ID. Then, user authentication and
app authorization is performed, by involving the client and Facebook
only. User authentication is done by logging in on Facebook (if necessary) and
authorization of the application is done by asking the user if he/she
wants to grant the application access to some data specified by the
application with the scope parameter. One should note, that \annSBW{ loose the comma} no
login data are sent to the application, thus, \annSBW{thus triffts mal garnicht, das komma is zu viel am besten ganzen satz anders} this is how Facebook
establishes a trust relation. 
If the user authorizes the application, an \textit{authorization code} is returned and
the user is redirected to the application, to which the
\textit{authorization code} is forwarded. During registration of the
application, an \textit{application secret} code was
created. In order to authenticate the application, the
\textit{authorization code} and the \textit{application secret} have
to be sent to Facebook's \textit{Graph API Token Endpoint}. Facebook
then returns an \textit{access token}, which can be included in the
URL by which the Graph API can be accessed. However, Facebook's
\ac{sdks} also provide functions to perform calls to the Graph API,
and in this case, the access token is included automatically.

\annSBW{HTTPS? Security überlegungen?}

\subsection{Graphs with DOT language}
\label{sec:graph-repr-with}
The \ac{dot} language allows for plain text description of
graphs. There are few programs that can process \ac{dot} graphs, most \annSBW{ not by a long shot}
of them are part of the
\textit{Graphviz}\footnote{http://www.graphviz.org}-package. 
The language allows for directed and undirected graphs, attributes can
be set and colors as well as different shapes for nodes and edges can
be used. The basic syntax is very easy, the graph shown in figure
\ref{fig:dot} has the \ac{dot}-code given in listing \ref{list:dot}.

@@

\subsection{Software Metrics}
\label{sec:metrics}
Metrics define a mapping of a `particular measurable entity to a
numerical value' \cite{Lanz06a}. Thus, they allow for analysis and summarization
of properties which are oftentimes not measurable\annSBW{gerad eben haste noch das gegenteil behauptet}. In addition, they
facilitate detection of `outliers in large amounts of data'
\cite{Lanz06a}. Metrics can be plotted as diagrams or charts, or they
can be used for visualizations such as graphs connecting nodes
characterized by metrics. Since humans are trained in understanding signs and
pictures, visualizations supports understanding and identifying of
`hidden aspects'. \cite{Lanz06a} In this paper, they are
employed to characterize users of the quiz. Some metrics can be output
as \ac{pdf}, and others are used to create the graphs in \ac{dot} language.

@@

\subsubsection{About The Quiz}
\label{sec:aboutQuiz}
The quiz is about dental health fact and also challenging\annSBW{loose that}. It can be
found at the URL \url{http://apps.facebook.com/zahniquiz}.
Each participants is asked five questions randomly selected
from a set of twenty questions. A question has four answers, of
which either one, two, three or each answer can be correct, thus
the kind of answers refers to the term \textit{multiple selection} \annSBW{der Satz is ganz im Arsch}.
A question is scored to be answered correctly, if the correct subset
was selected by the user, thus \annSBW{i.E. ... thus should be used seldom if at all} each correct and no incorrect answer
were selected. Each question answered correctly scores one, thus the
highest score is five. Figure \ref{fig:quiz} gives a screenshot of the quiz.

@@

\subsubsection{Implementation}
\label{sec:implementation}
The studied application makes use of 
Facebook's php-\ac{sdk} and JavaScript-\ac{sdk}\footnote{http://developers.facebook.com/docs/}. It requires
permissions (see \ref{sec:auth-auth}) for \textit{user$\_$birthday} and
\textit{publish$\_$stream} \annSBW{erklären!}.
The questions and answers of the quiz are stored in a MySQL-database. PHP scripts
handle the creation of the quiz and
the Pear DB module \footnote{http://pear.php.net/} is used
for database queries.

@@

\paragraph{Application flow}
\label{sec:application-flow}
When a user accesses the canvas page of the quiz, 
the \textit{OAuth Dialog} is started automatically by the
\textit{controller} \annSBW{vorhin haste noch gesagt FB macht das}. If authorization and authentication are completed
successfully, the user's basic data as well as the birthday are stored together with
a timestamp in the \textit{User} table. Column \textit{has-used-app} is set to 1. In
addition, each user's friends list is also stored in \textup{User} and
\textit{Friends} table, and for these \textit{users}, columns
\textit{has-used-app} is set to 0.

% the quiz is created by randomly choosing
% five questions from the questions table and requiring the answers. To
% make the questions more attractive, the name of an image convenient to
% the questions topic is also stored in the questions table, and this
% image is placed on the left side of each question. Since multiple
% selection is possible, before the answers, which are also queried from
% the MySQL database, checkboxes are placed before each answer. 
Next, the quiz is created as mentioned in section
\ref{sec:implmenetation}.
After completion, its is evaluated the final score is computed. As
each answer ought to be stored and multiple selection of answers is \annSBW{multiple answers are}
possible, user answers have to be stored in  a multidimensional \annSBW{what?}.

@@

\subsection{Graphs detailing the spreading of the quiz}
\label{sec:distribution-quiz}
In this section, graphs illustrating how the
application spread \annSBW{-s/-ed} are presented. The original user names collected
from Facebook were made anonymous
 (see section \ref{sec:data-evaluation}). Friendships are
represented by directed edges: If user A has user B in the friend list, an
arrow exists targeting user B.
